Communicating with convexity

### 2017-11-23 article_cvxpy.tex
water-filling plot updated

% edited Keith Briggs 2017-11-16
% latexmk -pdf article_cvxpy_2017-11-16 && e article_cvxpy_2017-11-16.pdf &

@@

 One of the simplest problems to solve in communications with convex optimisation is the classic \textit{water-filling} problem.  This is where a total power $P$ is to be assigned to $n$ different communications channels, with the objective of maximising the total communication rate.  The communication rate of the $i$th channel, $r_i$ is given by \citep[p.245]{cvxpybook}, 
 \begin{align}
  r_i = \log(x_i+\alpha_i),
 \end{align}
 where $x_i$ is the power allocated to the $i$th channel and $\alpha_i$ is the floor above the baseline at which power can be added to the channel.  This can be seen in Figure \ref{fig:waterfilling} as being analogous to filling an uneven basin to a constant level (hence the name \textit{water-filling}).

 Noting that all power allocations must be non-negative, the optimisation problem can therefore be formulated as,
 \begin{align}
\text{minimise} \quad  -&\sum_{i=1}^{n} \log(x_i+\alpha_i) \nonumber \\
\text{subject to} \quad &x_i \geq 0, \quad \forall i \nonumber \\
&\sum_{i=1}^{n}x_i = P.
 \end{align}
 This is a convex problem since $-\log(x_i+\alpha_i)$ has positive curvature within the domain $x_i \geq 0$, and all the constraints are affine and therefore convex.

\subsection{Implementation}
We decided to solve the water-filling problem using \texttt{cvxpy} with $n=3$ channels and $(\alpha_1,\alpha_2,\alpha_3) = (0.8,1.0,1.2)$.  For simplicity we chose a dimensionless total power $P=1$.  Using these parameters, the maximum communication rate is, $\sum_{i=1}^{n}r_i = 0.863$, which is achieved with power allocations $(x_1,x_2,x_3) = (8/15,5/15,2/15)$.  This is the solution shown in Figure \ref{fig:waterfilling}.

@@

\begin{figure}[H]
\centering
\includegraphics[width=1.0\linewidth]{water_filling_plot.pdf}
\caption{The greatest amount of power, $x$, is allocated to the channel with the lowest $\alpha$.}
\label{fig:waterfilling}
\end{figure}

@@

\subsection{Implementation}
For ease of terminology, we will make a few simplifying assumptions.  First, we set the number of transmitters equal to the number of receivers, so $n=m$.  We also pair receiver $i$ to transmitter $k$ such that $i=k$.  This means that the recipient matrix $\Delta$ is the $n\times n$ identity matrix.

@@

To generate the gain matrix $G$, we randomly sample distances from a given distribution.  For paired transmitters and receivers ($k=i$), we sample $d$ from a scaled $\beta(2,2)$ distribution, such that $d_{ii} \sim 0.2\times\text{Beta}(2,2)$.  For unpaired transmitters and receivers ($k\neq i$), we sample $d$ from a uniform distribution, such that $d_{ik} \sim \text{Uniform}(0,1)$.  This is illustrated in Figure \ref{fig:dist}.

@@

\subsection{Example code}
For the random seed of 100 given, the optimal total power consumption is $P = 0.381$, with $p_1,p_2,p_3 = (0.139, 0.072, 0.169)$.  The time taken to find this solution was around 3ms.  Around 70\% of the simulations with these parameters have a feasible solution, with an average optimal power consumption of $P\sim0.4$.

>>>>

% edited Keith Briggs 2017-11-23
% edited Keith Briggs 2017-11-16
% latexmk -pdf article_cvxpy && e article_cvxpy.pdf &

@@

\noindent{}One of the simplest problems to solve in communications with convex optimisation is the classic \textit{water-filling} problem.  Here a total power $P$ is to be assigned to $n$ different communications channels, with the objective of maximising the total communication rate.  By Shannon's theorem \citep[p.245]{cvxpybook}, an upper bound for the rate is given by $\log(1+x_i/\alpha_i)$, where $x_i$ be the power allocated to the $i$th channel and $\alpha_i$ is the noise power in that channel. This can be written as $\log(\alpha_i+x_i)-\log(\alpha_i)$ and the constants $\log(\alpha_i)$ disregarded for the purposes of maximization, and thus the communication rate $r_i$ of the $i$th channel is given by
 \begin{align}
  r_i = \log(x_i+\alpha_i).
 \end{align}
This can be seen in Figure~\ref{fig:waterfilling} as being analogous to filling an uneven basin to a constant level (hence the name \textit{water-filling}).

Noting that all power allocations must be non-negative, the optimisation problem can therefore be formulated as,
 \begin{align}
\text{minimise} \quad  -&\sum_{i=1}^{n} \log(x_i+\alpha_i) \nonumber \\
\text{subject to} \quad &x_i \geqslant 0, \quad \forall i \nonumber \\
&\sum_{i=1}^{n}x_i = P.
 \end{align}
This is a convex problem since $-\log(x_i+\alpha_i)$ has positive curvature within the domain $x_i \geqslant 0$, and all the constraints are affine and therefore convex.

\subsection{Implementation}
\noindent We decided to solve the water-filling problem using \texttt{cvxpy} with $n=3$ channels and $(\alpha_1,\alpha_2,\alpha_3) = (0.8,1.0,1.2)$.  For simplicity we chose a dimensionless total power $P=1$.  Using these parameters, the maximum communication rate is, $\sum_{i=1}^{n}r_i = 0.863$, which is achieved with power allocations $(x_1,x_2,x_3) = (8/15,5/15,2/15)$.  This is the solution shown in Figure~\ref{fig:waterfilling}.

@@

\begin{figure}[H]
\centering
\includegraphics[width=1.0\linewidth]{water_filling_plot.pdf}
\caption{The optimal solution to the water-filling problem allocates the greatest amount of power, $x$, to the channel with the lowest noise power, $\alpha$.}
\label{fig:waterfilling}
\end{figure}

@@

\subsection{Implementation}
\noindent For ease of terminology, we will make a few simplifying assumptions.  First, we set the number of transmitters equal to the number of receivers, so $n=m$.  We also pair receiver $i$ to transmitter $k$ such that $i=k$.  This means that the recipient matrix $\Delta$ is the $n\times n$ identity matrix.

@@

To generate the gain matrix $G$, we randomly sample distances from a given distribution.  For paired transmitters and receivers ($k=i$), we sample $d$ from a scaled $\beta(2,2)$ distribution, such that $d_{ii} \sim 0.2\times\text{Beta}(2,2)$.  For unpaired transmitters and receivers ($k\neq i$), we sample $d$ from a uniform distribution, such that $d_{ik} \sim \text{Uniform}(0,1)$.  This is illustrated in Figure~\ref{fig:dist}.

@@

\subsection{Example code}
\noindent For the random seed of 100 given, the optimal total power consumption is $P = 0.381$, with $p_1,p_2,p_3 = (0.139, 0.072, 0.169)$.  The time taken to find this solution was around 3ms.  Around 70\% of the simulations with these parameters have a feasible solution, with an average optimal power consumption of $P\sim0.4$.
